Reintroduction of resistant frogs facilitates landscape-scale recovery in the presence of a lethal fungal disease

### 2024-05-21 doc/manuscript/cover_letter_nature_comm.qmd
Final edits prior to submission

---
format: 
  pdf:
    fig-pos: "H"
    include-in-header: 
      text: '\pagenumbering{gobble}'
editor: visual
editor_options:
  markdown:
    wrap: sentence
---

In this cover letter, I attempt to provide some context for our manuscript, "Reintroduction of resistant frogs facilitates landscape-scale recovery in the presence of a lethal fungal disease".
It is my sincere hope that this context will serve as a more compelling argument for why this manuscript is worthy of consideration for publication in *Nature Communications* than the usual (and somewhat pedantic) arguments of how novel this research is or how important it is for a particular branch of science.

@@

During my own 30+ year career studying mountain ecosystems and their fauna, I've experienced those disappearances first hand and more times than I care to remember.
After restoring lakes to their original fishless condition and documenting the subsequent recolonization and skyrocketing abundance of Sierra Nevada yellow-legged frogs, I've watched helplessly as those same amphibians succumbed to Bd and their mountain home went silent.
Gone was the shower of frogs diving from the shore into water as they fled my approach.
Gone were the masses of tadpoles in lake shallows, once so dense that the water frothed with their wriggling tails.
Diminished too were the birds and snakes that fed on those frogs.
These protected habitats in Yosemite National Park were unraveling like a tapestry come unknotted.

Unwilling to simply walk away from this loss, my colleagues and I put everything we had into understanding this pathogen and its impact on amphibians.
We gained key insights from our decades of research and from sharing results with others around the world who were witnessing their own amphibian catastrophes.
The work was slow and arduous, with as many setbacks as successes, and all the while the threat of extinction loomed large.
A key inflection point for me came twenty years ago with the realization that some frog populations driven to near extinction by Bd were, almost miraculously, beginning to recover on their own.
The subsequent intensive study of those populations provided the knowledge that we needed to bolster this natural recovery using frog reintroductions.

The science behind that uniquely successful 15 year recovery effort is the focus of our manuscript, a story told through graphs, tables, and methodological minutiae.
But there is another way to understand the consequences of that work, and that is to sit on the shore of any one our study lakes in which we reintroduced frogs 10 or more years ago.
The silence that settled over the lake like a morning fog following the frogs' extinction is no more.
Instead, the lake shore is alive with frogs, one atop every rock and lining the sandy beaches.
On calm mornings, pairs of eyes bulge above the water's surface like unblinking periscopes.
The tadpoles too are back, their black thousands filling bays with ripples and splashing.
The lake is alive again.

>>>>

---
format: 
  pdf:
    fontsize: 12pt
    include-in-header: 
      text: '\pagenumbering{gobble}'
editor: visual
editor_options:
  markdown:
    wrap: sentence
---

In this cover letter, I'd like to provide some context for our manuscript, "*Reintroduction of resistant frogs facilitates landscape-scale recovery in the presence of a lethal fungal disease*".
Although this cover letter takes a different approach than most cover letters (including those I have written for other papers), it is my sincere hope that the provided context will serve as a compelling argument for why this manuscript is worthy of consideration for publication in *Nature Communications*.

@@

During my own 30+ year career studying mountain ecosystems and their fauna, I've experienced those disappearances first hand and more times than I care to remember.
After restoring lakes to their original fishless condition and documenting the subsequent recolonization and skyrocketing abundance of Sierra Nevada yellow-legged frogs, I've watched helplessly as those same amphibians succumbed to Bd, and their mountain home went silent.
Gone was the shower of frogs diving from the shore into water as they fled my approach.
Gone were the masses of tadpoles in lake shallows, once so dense that the water frothed with their wriggling tails.
Diminished too were the birds and snakes that fed on those frogs.
These protected habitats in Yosemite National Park were unraveling like a tapestry come unknotted.

Unwilling to simply walk away from this loss, my colleagues and I put everything we had into understanding Bd and its impact on amphibians.
We gained key insights from our decades of research and from sharing results with others around the world who were witnessing their own amphibian catastrophes.
The work was slow and arduous, with as many setbacks as successes, and all the while the threat of extinction loomed large.
A key inflection point for me came twenty years ago with the realization that some frog populations driven to near extinction by Bd were, almost miraculously, beginning to recover on their own.
The subsequent intensive study of those populations provided the knowledge that we needed to bolster this natural recovery using frog reintroductions.

The science behind that uniquely successful 15 year recovery effort is the focus of our manuscript, a story told through graphs, tables, and methodological minutiae.
But there is another way to understand that work, and that is to sit on the shore of any one of our study lakes in which we reintroduced frogs ten or more years ago.
The silence that settled over the lake like a morning fog following the frogs' extinction is no more.
Instead, the lake shore is alive with frogs, atop rocks and lining the sandy beaches.
On calm mornings, pairs of eyes bulge above the water's surface like unblinking periscopes.
The tadpoles too are back, their black thousands filling bays with ripples and splashing.
The lake is alive again.

### 2024-05-21 doc/manuscript/supplementary_information.qmd
Final edits prior to submission

---
format: 
  pdf:
    fig-pos: "H"
#format:
#   latex: 
#      cite-method: natbib
include-in-header:
  text: |
    \usepackage{blkarray}
include-before-body: 
  text: |
    \captionsetup{format=plain, labelfont=bf, labelsep=period}
editor: visual
editor_options:
  markdown:
    wrap: sentence
crossref: 
  fig-title: "**Supplementary Fig.**"
  fig-prefix: "Supplementary Fig."
  tbl-title: "**Supplementary Table**"
  tbl-prefix: "Supplementary Table"
bibliography: translocation.bib
csl: nature.csl
link-bibliography: false
---

@@

| **Parameter**                                                             | **Value**            | **Source**                                                                        |
|:--------------------------------------------------------------------------|:---------------------|:----------------------------------------------------------------------------------|
| $\sigma_{L_1}$, Yearly survival probability of year-1 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge                |
| $\sigma_{L_2}$, Yearly survival probability of year-2 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge                |
| $\sigma_{L_3}$, Yearly survival probability of year-3 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge                |
| $\sigma_{J_1}$, Yearly survival probability of year-1 juveniles           | Varies yearly        | Varies. Bounds estimated from field data, observations, natural history knowledge |
| $\sigma_{J_2}$, Yearly survival probability of year-2 juveniles           | 0.5                  | Estimated from field data, observations, natural history knowledge                |
| $\sigma_{A_R}$, Yearly survival probability of naturally recruited adults | Varies by population | Estimated from CMR studies                                                        |
| $\sigma_{A_T}$, Yearly survival probability of translocated adults        | Varies by population | Estimated from CMR studies                                                        |
| $p_{L_1}$, Probability of a year-1 tadpoles remaining as a tadpoles       | 1                    | Estimated from field data, observations, natural history knowledge                |
| $p_{L_2}$, Probability of a year-2 tadpoles remaining as a tadpoles       | 0.25                 | Estimated from field data, observations, natural history knowledge                |
| $p_{J_1}$, Probability of a year-1 juvenile remaining as a juvenile       | 0.25                 | Estimated from field data, observations, natural history knowledge                |
| $p_F$, Probability of a adult female reproducing in a year                | 0.5                  | Could be as high at 1, based on field observations                                |
| $F$, Number of surviving eggs produced by an adult female                 | 100                  | From observations of captive frogs                                                |

>>>>

---
format: 
  pdf:
    fig-pos: "H"
#format:
#   latex: 
#      cite-method: natbib
include-in-header:
  text: |
    \usepackage{blkarray}
    \usepackage{lineno}\linenumbers
include-before-body: 
  text: |
    \captionsetup{format=plain, labelfont=bf, labelsep=period}
editor: visual
editor_options:
  markdown:
    wrap: sentence
crossref: 
  fig-title: "**Supplementary Fig.**"
  fig-prefix: "Supplementary Fig."
  tbl-title: "**Supplementary Table**"
  tbl-prefix: "Supplementary Table"
bibliography: translocation.bib
csl: nature.csl
link-bibliography: false
---

@@

| **Parameter**                                                             | **Value**            | **Source**                                                                        |
|:--------------------------|:---------------|:-----------------------------|
| $\sigma_{L_1}$, Yearly survival probability of year-1 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge                |
| $\sigma_{L_2}$, Yearly survival probability of year-2 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge                |
| $\sigma_{L_3}$, Yearly survival probability of year-3 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge                |
| $\sigma_{J_1}$, Yearly survival probability of year-1 juveniles           | Varies yearly        | Varies. Bounds estimated from field data, observations, natural history knowledge |
| $\sigma_{J_2}$, Yearly survival probability of year-2 juveniles           | 0.5                  | Estimated from field data, observations, natural history knowledge                |
| $\sigma_{A_R}$, Yearly survival probability of naturally recruited adults | Varies by population | Estimated from CMR studies                                                        |
| $\sigma_{A_T}$, Yearly survival probability of translocated adults        | Varies by population | Estimated from CMR studies                                                        |
| $p_{L_1}$, Probability of a year-1 tadpoles remaining as a tadpoles       | 1                    | Estimated from field data, observations, natural history knowledge                |
| $p_{L_2}$, Probability of a year-2 tadpoles remaining as a tadpoles       | 0.25                 | Estimated from field data, observations, natural history knowledge                |
| $p_{J_1}$, Probability of a year-1 juvenile remaining as a juvenile       | 0.25                 | Estimated from field data, observations, natural history knowledge                |
| $p_F$, Probability of a adult female reproducing in a year                | 0.5                  | Could be as high at 1, based on field observations                                |
| $F$, Number of surviving eggs produced by an adult female                 | 100                  | From observations of captive frogs                                                |
